Embeddings and Attention in Predictive Modeling

### 2022-03-29 manuscript/manuscript.Rmd
move pca/tsne technical stuff to appendix

For Model 4, the neural network with multi-dimensional embeddings, we would need to treat the learned values before we can plot them. Specifically, we need to reduce the dimension of the data to 2D for embeddings with more than two dimensions. The standard approachs for this are principal component analysis (PCA; @shlens2014tutorial) and t-distributed stochastic neighbor embedding (t-SNE; @Maaten2008a), both of which we illustrate.

### Principal Component Analysis

@@

### t-Distributed Stochastic Neighbor Embedding

While t-SNE is a powerful visualization technique, there are a few points the modeler should be mindful of in practice. We refer the reader to @wattenberg2016how for a more thorough discussion, but stress here that the resulting plot can be very sensitive to the perplexity hyperparameter. In addition, for our use case, which is to explore the learned embeddings, we have fewer data points than most other applications in the literature. Accordingly, our reasonable range of hyperparameter values need to be adjusted down. In Figures \@ref(fig:flood-zone-tsne), we display the t-SNE of the learned embeddings for `flood_zone` in Model 3 with various perplexity settings: 2, 3, 5, and 10; the learning rate and the number of steps are fixed at 100 and 10,000, respectively, for all plots. In the first three plots, we can see the semblance of clusters forming. In particular, flood zone D appears in its own cluster, which implies it is far away from other levels in the embedding space, *as seen by the model*. When the perplexity is set to 10 (and beyond), however, we start to lose the geometric properties.

@@

>>>>

For Model 4, the neural network with multi-dimensional embeddings, we would need to treat the learned values before we can plot them. Specifically, we need to reduce the dimension of the data to 2D for embeddings with more than two dimensions. The standard approaches for this are principal component analysis (PCA; @shlens2014tutorial) and t-distributed stochastic neighbor embedding (t-SNE; @Maaten2008a), both of which we illustrate. For a technical review of these methods, refer to Appendix \ref{dim-red].

@@

In Figures \@ref(fig:flood-zone-tsne), we display the t-SNE of the learned embeddings for `flood_zone` in Model 3 with various perplexity settings: 2, 3, 5, and 10; the learning rate and the number of steps are fixed at 100 and 10,000, respectively, for all plots. In the first three plots, we can see the semblance of clusters forming. In particular, flood zone D appears in its own cluster, which implies it is far away from other levels in the embedding space, *as seen by the model*. When the perplexity is set to 10 (and beyond), however, we start to lose the geometric properties. While t-SNE is a powerful visualization technique, there are a few points the modeler should be mindful of in practice. We refer the reader to @wattenberg2016how for a more thorough discussion, but stress here that the resulting plot can be very sensitive to the perplexity hyperparameter. In addition, for our use case, which is to explore the learned embeddings, we have fewer data points than most other applications in the literature. Accordingly, our reasonable range of hyperparameter values are adjusted down.

@@

# Dimensionality Reduction Techniques {#dim-red}

## Principal Component Analysis

## t-Distributed Stochastic Neighbor Embedding
